---
title: "Data Exploration Assignment"
author: "Lorenzo Powell"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Since the release of the College Scorecard in September 2015 student enrollment has been directed towards high-earning institutions. My lack of skill with R has prevented me from being able to clearly demonstrate this graphically and with a cleaned regression model. This will be amended in a future commit; at present a series of multivariate regression models to support this hypothesis have been provided with adequate controls, high degree of statistical significance across models, and different regression types with a consistent objective. 


Load the necessary packages 
```{r packages, echo = FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(fixest)
library(vtable)
library(Ecdat)
library(wooldridge)
library(ggplot2)
library(ggstance)
library(NHANES)
library(dagitty)
library(dplyr)
library(purrr)
library(readr)
library(stringr)
library(lubridate)

```

Load the data files from github repository
```{r Load Data, echo = FALSE, warning=FALSE, error=FALSE, collapse=TRUE}

scorecard <-scorecard <- read_csv('../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/Most+Recent+Cohorts+(Scorecard+Elements).csv') %>%
  rename(opeid = OPEID) %>%
  rename(unitid = UNITID) %>%
  rename(md_earn_wne_p10 = 'md_earn_wne_p10-REPORTED-EARNINGS')
  
name_link <- read_csv('../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/id_name_link.csv')  %>%
  group_by(schname) %>%
  mutate(n = n()) %>%
  filter(n == 1)

TUT_files <- list.files(path = '../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/',
                       pattern = 'trends_up_to_', full.names = TRUE)

TUT_map <- map_df(TUT_files, read_csv)

```


The following served as my base mutations of the provided data sets. Very little was changed, the goal was to use these data-frames as references during my analysis.

```{r Mutations, echo = TRUE, warning=FALSE, error=FALSE}
# Trends Up TO (TUT) files were compiled in TUT_map // this mutation only serves to ease sorting 
TUT_mutate <- TUT_map %>%
    mutate(start_mo_week = ymd(str_sub(monthorweek, start = 0, end = 10))) %>%
    group_by(schname, keyword)

# Mean & SD of all index's
index_mean <- mean(as.numeric(TUT_mutate$index), na.rm = TRUE)
index_sd <- sd(as.numeric(TUT_mutate$index), na.rm = TRUE)

# Standardized TUT files
TUT_standard <- TUT_mutate %>%
  group_by(schname, keyword) %>%
  mutate(index_stand = ((index - index_mean)/index_sd)) %>%
  mutate(start_year = floor_date(start_mo_week, 'year'))

# Merge of all exploratory data sets including standardized values
big_merge_stand <- inner_join(name_link, TUT_standard, by = 'schname')
big_merge_stand <- inner_join(big_merge_stand, scorecard, by = 'opeid')

# A mutated merge with percentile data grouping 10 year income of graduates into classes 'Low', 'Med', 'High'
perc_merge <- big_merge_stand %>%
  mutate(inc_group = case_when(
    between(as.numeric(md_earn_wne_p10), 0, 31100) ~ "Low",
    between(as.numeric(md_earn_wne_p10), 31101, 44600) ~ "Med",
    between(as.numeric(md_earn_wne_p10), 44601, 166200) ~ "High",
    FALSE ~ NA_character_)) %>%
  group_by(schname, keyword, inc_group)

# Annualized Merge Data 
merge2013 <- perc_merge %>%
  filter(start_year == '2013-01-01')
merge2014 <- perc_merge %>%
  filter(start_year == '2014-01-01')
merge2015 <- perc_merge %>%
  filter(start_year == '2015-01-01')
merge2016 <- perc_merge %>%
  filter(start_year == '2016-01-01')
```

10 year median earnings can be identified by the three following groups:

  Low:    0 -31100
  Med:    31101 - 44600
  High:   44601- 166200 
  
Though only average values in 3 tiers were calculated there is evidence of left skewed distribution, a T-test would be useful for small sub-sample regression to properly consider the impact of high earning outliers. 
```{r Earnings Percentiles, echo = TRUE, warning=FALSE, error=FALSE}
# Use percentile values to compare index rends
quantile_dat <- as.numeric(big_merge_stand$md_earn_wne_p10)
quant_dat <- na.omit(quantile_dat)
quantile(quant_dat, c( 0.25, 0.75, 1))

low_inc <- (0:31100)
med_inc <- (31101:44600)
high_inc <- (44601:166200)

# Rough distribution plot of 10 year earnings associated with index scores
quant_plot <- plot(quantile(quant_dat, c( 0.25, 0.75, 1 )))
('../su-omsba-data_exploration_assignment-lpowell/quant_plot.png')
```

Over time, the average 10-year earnings after enrollment is expected to change by -0.037 units; however, when controlling for the standardized-index value there will be a 0.4274 unit change for a 1 unit increase in enrollment year. When controlling for standardized-index scores, our hypothesis supported at the 5% significance level. These statistics provide insight on the expected relationship we should expect to see between overall enrollment volume and and index scores; if enrollment period did not have a notable association with enrollment volume, then 10 year median earnings and index scores should be neutral or unrelated to enrollment.
```{r Enrollment Year & Earnings, echo = TRUE, warning=FALSE, error=FALSE}
# Year o enrollment has a significant negative association with 10 year earnings // we see the opposite relationship in the setting of index rating of 
year_start_earn <- feols(as.numeric(md_earn_wne_p10) ~ start_year, data = big_merge_stand)
year_start_earn_con_ind <- feols(as.numeric(md_earn_wne_p10) ~ start_year | index_stand, data = big_merge_stand)

etable(year_start_earn, year_start_earn_con_ind)
summary(year_start_earn_con_ind)

```

The section below is an exploratory one demonstrating that greater '10 Year Earnings' tends to have a significant negative association with "Average Debt'. Because I did not manage to group earnings as 'Low', 'Med', or 'High', each earning group is included in the regression table.
```{r Debt & Earnings, echo = TRUE, warning=FALSE, error=FALSE}
# Median debt associated with earnings
med_earn_debt_school_df <- distinct(select(big_merge_stand, schname, md_earn_wne_p10, GRAD_DEBT_MDN_SUPP))
med_earn_debt_school_fe_reg <- feols(as.numeric(GRAD_DEBT_MDN_SUPP) ~ md_earn_wne_p10 | schname, data = big_merge_stand)
med_earn_debt_school_fe_reg2 <- feols(as.numeric(GRAD_DEBT_MDN_SUPP) ~ md_earn_wne_p10 + start_year| index_stand, data = big_merge_stand)
med_earn_debt_school_fe_reg_het <- feols(as.numeric(GRAD_DEBT_MDN_SUPP) ~ md_earn_wne_p10 + start_year| index_stand, data = big_merge_stand, vcov = 'hetero')
```

The fixed-effects model 'index_enroll_reg' showed to have the best predictive properties of my multivariate regression models. Its interpretation in the setting of observed fixed variale effects of '10 Year Earnings', 'Post-Grad Debt', and 'Institution Locale', a 1 unit change in the 'Enrollment Period' is associated with a -0.0005 unit change in the 'Standardized Index Score'.
Another notable model which controls for 'Institution Name' to hopefully avoid an unobserved effect of institution prestige. 'date/year_index_reg' differ in enrollment period level from month & week to annual intervals. These model's do not account for the aforementioned post-graduate financial outcomes but have comparable R2 values; however these models suggest a strong positive relationship between 'Enrollment Period' and 'Standardized Index Score'.
Had I managed to apporpriately group the data by quantile as calculated above, Dummy variables could have been used in either set of model's to avoid the slough of values in the regression table like that of 'med_earn_debt_school_fe_reg' and its alternates.
```{r Standardized Index Ratings & Enrollment Period, echo = TRUE, warning=FALSE, error=FALSE}
# Index by time
date_index <- select(big_merge_stand, schname, schid, start_mo_week, start_year, index, index_stand) %>%
  mutate(start_year = floor_date(start_mo_week, 'year')) %>%
  group_by(start_year)
date_index2013 <- date_index %>%
  filter(start_year == '2013-01-01') #%>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2014 <- date_index %>%
  filter(start_year == '2014-01-01')# %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2015 <- date_index %>%
  filter(start_year == '2015-01-01') # %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2016 <- date_index %>%
  filter(start_year == '2016-01-01')# %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()

# The effects of time on standardized index scores in both an annual and month/week breakout
date_index_reg <- feols(index_stand ~ start_mo_week + schid, data = date_index)
year_index_reg <- feols(index_stand ~ start_year + schid, data = date_index)
etable(date_index_reg, year_index_reg)

# The effects of school selection on standardized index scores
school_index_fe_reg <- feols(index_stand ~ schid | start_year, data = date_index)
school_index_reg <- feols(index_stand ~ schid + start_year, data = date_index)
etable(school_index_reg, school_index_fe_reg)

# Standardized index associated with enrollment period  //  includes FE's fpr 10 year earnings and median debt
index_enroll_reg <- feols(index_stand ~ start_mo_week| md_earn_wne_p10 + GRAD_DEBT_MDN_SUPP + LOCALE, data = big_merge_stand)
etable(index_enroll_reg)
```

When running regression similar regression model's with 'Earnings Class' but focused on institution 'schid' choice, it can be understood that a 1 unit increase in the 'Enrollment Period' is associated with reduced likelihood (-0.102 units) to enroll at that institution; interestingly the location of the institution was also shown to be significant but its impact on the overall model was negligible. Adding the fixed effect variable for 'Average Debt' also reduced reliability, my exploratory research showed that both '10 Year Earnings' and 'Average Debt' indicators of 'Earnings Class'; though not an outright hindrance, removal of debt from the fixed effects model added marginal improvement in statistical significance whereas its presence would have provided incremental improvements to within group residuals (within R2).
```{r Institution & Earning Class, echo = TRUE, warning=FALSE, error=FALSE}
# Median debt associated with earnings
school_enroll_inc <- feols(schid ~ start_mo_week + index_stand + inc_group, data = perc_merge)
school_enroll_inc_fe <- feols(schid ~ start_mo_week | index_stand + inc_group, data = perc_merge)
etable(school_enroll_inc, school_enroll_inc_fe)
```

None of my attempts at creating a useful graph were successful. Most of my time has been spent trying to create appropriate variable levels. Some of the remaining scrap lines can be reviewed in in 'DAE_explore_data.R'. Cluttered data and long run-time's made it difficult to generate a clean visualization. The second graphic which compares 'Standardized Index Scores' grouped by 10 year 'Earnings Class' can be interpreted by looking at the slope of the observed line from the year start. There is a clear graphical correlation between 'Enrollment Period' and Index score, looking at the observation count over time by 'Earnings Class' also indicates growth in enrollment volume. 
```{r Graphics, echo = TRUE, warning=FALSE, error=FALSE}
# Runs as apotential base but is otherwise useless
date_index2013 <- date_index %>%
  filter(start_year == '2013-01-01') 
date_index2014 <- date_index %>%
  filter(start_year == '2014-01-01')
date_index2015 <- date_index %>%
  filter(start_year == '2015-01-01')
date_index2016 <- date_index %>%
  filter(start_year == '2016-01-01')
ggplot() + 
  geom_line(data = date_index2013, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2013'))+
  geom_line(data = date_index2014, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2014'))+
  geom_line(data = date_index2015, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2015'))+
  geom_line(data = date_index2016, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2016'))

# An unclear visual representation of a clear increase in volume of high earning institutions // may be interpreted as increased attention
inc_group_low<- perc_merge %>%
  filter(inc_group == 'Low') 
inc_group_med<- perc_merge %>%
  filter(inc_group == 'Med')
inc_group_high<- perc_merge %>%
  filter(inc_group == 'High')

ggplot() + 
  geom_line(data = inc_group_low, mapping = aes(x = start_year, y = index_stand, color = 'low'))+
  geom_line(data = inc_group_med, mapping = aes(x = start_year, y = index_stand, color = 'med'))+
  geom_line(data = inc_group_high, mapping = aes(x = start_year, y = index_stand, color = 'high'))

```