---
title: "Data Exploration Assignment"
author: "Lorenzo Powell"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Since the release of the College Scorecard in September 2015 student enrollment has been directed towards high-earning institutions.


Load the necessary packages 
```{r packages, echo = FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(fixest)
library(vtable)
library(Ecdat)
library(wooldridge)
library(ggplot2)
library(ggstance)
library(NHANES)
library(dagitty)
library(dplyr)
library(purrr)
library(readr)
library(stringr)
library(lubridate)

```

Load the data files from github repository
```{r Load Data, echo = FALSE, warning=FALSE, error=FALSE, collapse=TRUE}

scorecard <-scorecard <- read_csv('../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/Most+Recent+Cohorts+(Scorecard+Elements).csv') %>%
  rename(opeid = OPEID) %>%
  rename(unitid = UNITID) %>%
  rename(md_earn_wne_p10 = 'md_earn_wne_p10-REPORTED-EARNINGS')
  
name_link <- read_csv('../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/id_name_link.csv')  %>%
  group_by(schname) %>%
  mutate(n = n()) %>%
  filter(n == 1)

TUT_files <- list.files(path = '../su-omsba-data_exploration_assignment-lpowell/data/Lab3_Rawdata/',
                       pattern = 'trends_up_to_', full.names = TRUE)

TUT_map <- map_df(TUT_files, read_csv)

```


The following served as my base mutations of the provided data sets. Very little was changed, the goal was to use these data-frames as references during my analysis.

```{r Mutations, echo = TRUE, warning=FALSE, error=FALSE}
# Trends Up TO (TUT) files were compiled in TUT_map // this mutation only serves to ease sorting 
TUT_mutate <- TUT_map %>%
    mutate(start_mo_week = ymd(str_sub(monthorweek, start = 0, end = 10))) %>%
    group_by(schname, keyword)

# Mean & SD of all index's
index_mean <- mean(as.numeric(TUT_mutate$index), na.rm = TRUE)
index_sd <- sd(as.numeric(TUT_mutate$index), na.rm = TRUE)

# Standardized TUT files
TUT_standard <- TUT_mutate %>%
  group_by(schname, keyword) %>%
  mutate(index_stand = ((index - index_mean)/index_sd)) %>%
  mutate(start_year = floor_date(start_mo_week, 'year'))

# Merge of all exploratory data sets including standardized values
big_merge_stand <- inner_join(name_link, TUT_standard, by = 'schname')
big_merge_stand <- inner_join(big_merge_stand, scorecard, by = 'opeid')

# Annualized Merge Data 
merge2013 <- big_merge_stand %>%
  filter(start_year == '2013-01-01')
merge2014 <- big_merge_stand %>%
  filter(start_year == '2014-01-01')
merge2015 <- big_merge_stand %>%
  filter(start_year == '2015-01-01')
merge2016 <- big_merge_stand %>%
  filter(start_year == '2016-01-01')

```


Over time, the average 10-year earnings after enrollment is expected to change by -0.037 units; however, when controlling for the standardized-index value there will be a 0.4274 unit change for a 1 unit increase in enrollment year. When controlling for standardized-index, our hypothesis supported at the 5% significance level.
```{r Enrollment Year & Earnings, echo = TRUE, warning=FALSE, error=FALSE}
# Year o enrollment has a significant negative association with 10 year earnings // we see the opposite relationship in the setting of index rating of 
year_start_earn <- feols(as.numeric(md_earn_wne_p10) ~ start_year, data = big_merge_stand)
year_start_earn_con_ind <- feols(as.numeric(md_earn_wne_p10) ~ start_year | index_stand, data = big_merge_stand)

etable(year_start_earn, year_start_earn_con_ind)
summary(year_start_earn_con_ind)

```


```{r Earnings & Debt, echo = TRUE, warning=FALSE, error=FALSE}
# MEDIAN EARNINGS ASSOCIATED W/ DEBT

med_earn_debt_school_df <- distinct(select(big_merge_stand, schname, md_earn_wne_p10, GRAD_DEBT_MDN_SUPP))
med_earn_debt_school_fe_reg <- feols(as.numeric(GRAD_DEBT_MDN_SUPP) ~ md_earn_wne_p10 | schname, data = big_merge_stand)

```


```{r Standardized Index Ratings & Enrollment Period, echo = TRUE, warning=FALSE, error=FALSE}
# Index by time
date_index <- select(big_merge_stand, schname, schid, start_mo_week, start_year, index, index_stand) %>%
  mutate(start_year = floor_date(start_mo_week, 'year')) %>%
  group_by(start_year)
date_index2013 <- date_index %>%
  filter(start_year == '2013-01-01') #%>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2014 <- date_index %>%
  filter(start_year == '2014-01-01')# %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2015 <- date_index %>%
  filter(start_year == '2015-01-01') # %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()
date_index2016 <- date_index %>%
  filter(start_year == '2016-01-01')# %>%
  #ggplot(aes(x = start_year, y = index-index_stand, color = start_year)) + geom_line()

# The effects of time on standardized index scores in both an annual and month/week breakout
date_index_reg <- feols(index_stand ~ start_mo_week + schid, data = date_index)
year_index_reg <- feols(index_stand ~ start_year + schid, data = date_index)
etable(date_index_reg, year_index_reg)

# The effects of school selection on standardized index scores
school_index_fe_reg <- feols(index_stand ~ schid | start_year, data = date_index)
school_index_reg <- feols(index_stand ~ schid + start_year, data = date_index)
etable(school_index_reg, school_index_fe_reg)


```

None of my attempts at creating a useful graph were successful. Most of my time has been spent trying to create appropriate variable levels. 
  
Load as PNG File 
```{r Graphics, echo = TRUE, warning=FALSE, error=FALSE}
# Runs as apotential base but is otherwise useless
date_index2013 <- date_index %>%
  filter(start_year == '2013-01-01') 
date_index2014 <- date_index %>%
  filter(start_year == '2014-01-01')
date_index2015 <- date_index %>%
  filter(start_year == '2015-01-01')
date_index2016 <- date_index %>%
  filter(start_year == '2016-01-01')
ggplot() + 
  geom_line(data = date_index2013, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2013'))+
  geom_line(data = date_index2014, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2014'))+
  geom_line(data = date_index2015, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2015'))+
  geom_line(data = date_index2016, mapping = aes(x = month(start_mo_week), y = mean(index_stand), color = '2016'))


```